---
title: "Data Integration"
author: "Hazim Fitri"
output: pdf_document
---

# Data Integration

## Import data in R

### Native data set

Using native data from R, we can call the data by using `data()`. For example:-

```{r}
data(iris)
```

If we wish to understand further regarding the data, we can put `?` in front of out datset. For example:-

```{r}
?iris
```

### Excel file (.xlsx)

In order to load .xlsx file, first we need to load library openxlsx. Then, we will be able to use function `read.xlsx()`. To see all the default value, run `?read.xlsx`

```{r}
library(openxlsx)
big_mart = read.xlsx('Big Mart Dataset.xlsx')
```

## Integrate Data

```{r}
mydata1 = read.table('mydata1.txt')
mydata2 = read.csv('mydata2.csv')
print(colnames(mydata1))
print(colnames(mydata2))
```

Both data has different column names. Thus, in order to integrate both data, we can use `cbind()` function which will combine both column into one data frame.

```{r}
mydata3 = cbind(mydata1, mydata2)
colnames(mydata3)
```

If there's any unique identifier, we can use `merge()` function and it'll automatically merge combine given datasets and merge column with the name name

```{r}
mydata4 = merge(mydata1, mydata2)
mydata4
```

However, if the column name is different for both data sets, we can specify by using parameter `by.x` and `by.y`

```{r}
mydata
```
