---
title: "Data Reduction"
author: "Hazim Fitri"
date: "2024-12-17"
output: pdf_document
---

# Data Reduction

Data reduction is a technique that we can use when the number of the data is too large and using full data requires a costly and time-consuming computational method. There are two techniques to reduce the data:

1)  Dimension Data Reduction

2)  Numerosity Data Reduction

## Dimensional Data Reduction

### Removing Attributes

```{r}
library(ISLR)
data(package='ISLR')
data("Hitters")
?Hitters
hitters2 = na.omit(Hitters)
model.f = lm(Salary~., data=hitters2)
summary(model.f)
```

As we can see from the above summary for linear model of salary with other variables, only certain variables can be considered significant as indicate by at least one '\*' at the right of the column. This indicates that the variables are at least significance at alpha =0.05

```{r}
hitters3 = cbind(hitters2$AtBat, hitters2$Hits, hitters2$Walks, hitters2$CWalks,
                 hitters2$Division, hitters2$PutOuts)
head(hitters3)
```

### Primary Component Analysis (PCA)

```{r}
reading = read.csv('READING120n.csv')
head(reading)
```

Remove non-numeric column.

```{r}
reading2 = reading[,-1]
head(reading2)
```

```{r}
library(psych)
describe(reading2)
```

```{r}
z = scale(reading2)
pairs(~., data=z)
```

```{r}
cor_z = cor(z)
cor_z
```

```{r}
eigen(cor_z)
```

```{r}
scree(cor_z)
```

```{r}
eigen(cor_z)$vectors
y = z %*% eigen(cor_z)$vectors
colnames(y) = c('PCA1', 'PCA2', 'PCA3', 'PCA4', 'PCA5', 'PCA6')
prop.var = eigen(cor_z)$values / length(eigen(cor_z)$values)
cumsum(prop.var)
```

### Factor Analysis

```{r}
fa = factanal(cor_z, factors=2, rotation='varimax')
?factanal
```
