---
title: "Data Integration"
author: "Hazim Fitri"
output: 
  pdf_document:
    toc: true
    toc_depth: 6
---

# Data Integration

## Import data in R

### Native data set

Using native data from R, we can call the data by using `data()`. For <example:->

```{r}
data(iris)
```

If we wish to understand further regarding the data, we can put `?` in front of out datset. For <example:->

```{r}
?iris
```

### Excel file (.xlsx)

In order to load .xlsx file, first we need to load library openxlsx. Then, we will be able to use function `read.xlsx()`. To see all the default value, run `?read.xlsx`

```{r}
library(openxlsx)
big_mart = read.xlsx('./Data/Big Mart Dataset.xlsx', sheet = 1, startRow = 1)
```

## Integration of data with different attributes.

```{r}
mydata1 = read.table('./Data/mydata1.txt')
mydata2 = read.csv('./Data/mydata2.csv')
print(colnames(mydata1))
print(colnames(mydata2))
```

Both data has different column names. Thus, in order to integrate both data, we can use `cbind()` function which will combine both column into one data frame.

```{r}
mydata3 = cbind(mydata1, mydata2)
colnames(mydata3)
```

If there's any unique identifier, we can use `merge()` function and it'll automatically merge combine given datasets and merge column with the same name. In this example both dataframe has column 'ID_Person'

```{r}
mydata4 = merge(mydata1, mydata2)
mydata4
```

However, if the column name is different for both data sets, we can specify by using parameter `by.x` and `by.y`

```{r}
load('./Data/mydata4.RData')
colnames(mydata4)
mydata5 = read.csv('./Data/mydata5.csv')
colnames(mydata5)
mydata6 = merge(mydata4, mydata5, by.x='ID', by.y='IDPerson')
mydata6
```

### Data integration based on inconsistent attribute names with some mismatched attribute values

It will remove all the row without matching value.

```{r}
mydata7 = mydata5[1:10, ]
mydata8 = merge(mydata4, mydata7, by.x='ID', by.y='IDPerson')
mydata8
```

However, we can still retain the unmatch data by adding argument `all=T`

```{r}
mydata9 = merge(mydata4, mydata7, by.x='ID', by.y='IDPerson', all=T)
colnames(mydata9)
```

Alternative way is we can rename the column before merge

```{r}
library(plyr)
mydata10 = rename(mydata9, c('ID'='Nombor ID', 'house'='Number of House'))
```

## Customize data attribute

### Edit manually

If we wish to customize our data, the most basic way to do it is by using `edit()` function to manually edit the value one by one.

```{r}
mydata11 = edit(mydata10)
```

Edit data

```{r}
datam1 = read.csv('./Data/dataM1.csv')
library(tidyverse)

```

```{r}
city_name = function(city) {
  city = tolower(city) #convert all letter to small letter
  city = trimws(city) #trim all white space
  city = gsub('+', '', city) #replace with 1 space
  city = tools::toTitleCase(city) #change format to title case
  return(city)
}

a = tolower(datam1$City)
b = trimws(a)
c = ?gsub('+', '', b)
d = tools::toTitleCase(c)
```

Edit inconsistent data

(E.g., shortform)

```{r}
datam2 = read.csv('./Data/dataM2.csv')
head(datam2)
```

First, look at the unique data of city column

```{r}
unique(datam2$City)
```

Map all short form to the full name. Then, create a function to convert the short form to the full name.

```{r}
city_map = list('NY'='New York', 'CHI'='Chicago', 'LA'='Los Angeles')
std_city = function(city){
  if(city%in%names(city_map)) {
    return(city_map[[city]])
  } else {
    return(city)
  }
}

datam2$City = sapply(datam2$City, std_city)

unique(datam2$City)
```

### Remove redundant data

```{r}
datam3 = read.csv('./Data/dataM3.csv',sep=';')
datam3
```

```{r}
library(dplyr)
duplicated(datam3) #check for duplicated data
distinct(datam3) #return only unique/distinct data
datam3 %>% distinct(datam3$id, .keep_all=T) #.keep_all is to keep all column
```

## Export from R

In order to export file from R, first thing that we need to know is where the will be exported to. To see the save filepath, we can use `getwd()` to see the directory

```{r}
getwd()
```

If we wish to change the save location, we can redefine our location by using setwd() function and put the path as argument

```{r}
getwd()
```

### Save .RData file

Now, we can proceed to save our very first file in our local storage. To save R file we can simply use:

```{r}
save(datam3, file='data_M3.RData')
```

### Save .csv file

```{r}
write.csv(datam3, file='data_M3.csv')
```

### Save .txt file

```{r}
write.table(datam3, file='data_M3.txt', sep='\t')
```
