---
title: "Tenaga Buruh"
output: 
  pdf_document:
    toc: true
    toc_depth: 6
geometry: margin=0.5cm
---

# Objektif Analisis

-   Meramal kadar pengangguran bagi 2 tahun akan datang

-   Analyze the impact of economic shocks on labor market recovery

-   Study the relationship between participation rates and unemployment

-   Identify seasonal employment patterns

-   Monitor long-term structural changes in the labor market

# Load library

```{r, warning=FALSE}
library(arrow)
library(forecast)
lf = read_parquet("https://storage.dosm.gov.my/labour/lfs_qtr.parquet")
```

# Tenaga Buruh

```{r}
lf_ts = ts(lf[,2], start = c(2010,1), frequency = 4)

plot.ts(lf_ts, main = 'Data Tenaga Buruh di Malaysia', xlab = 'Tahun', 
        ylab = 'Bilangan Tenaga Buruh')
```

## Decomposing labor force

```{r}
plot(decompose(lf_ts))
```

```{r}
plot.ts(decompose(lf_ts)$trend, 
        main = 'Trend Tenaga Buruh di Malaysia', 
        xlab = 'Tahun', ylab = 'Tenaga Buruh')
```

```{r}
plot.ts(decompose(lf_ts)$seasonal, 
        main = 'Corak Bermusim Tenaga Buruh di Malaysia', 
        xlab = 'Tahun', ylab = 'Tenaga Buruh', xlim = c(2010,2011))
```

```{r}
plot.ts(decompose(lf_ts)$random, 
        main = 'Corak Rawak Tenaga Buruh di Malaysia', 
        xlab = 'Tahun', ylab = 'Tenaga Buruh')
```

## Meramal Tenaga Buruh di Malaysia

```{r}
plot(forecast(auto.arima(lf_ts)), 
     main = 'Tenaga Buruh Malaysia: Ramalan 2025-2026', 
     xlab = 'Tahun', ylab = 'Bilangan Tenaga Buruh')
```

# Employed labor force

```{r}
lf_emp = ts(lf[,3], start = c(2010,1), frequency = 4)

plot.ts(lf_emp, main = 'Tenaga Buruh Bekerja di Malaysia',
        xlab = 'Tahun', ylab = 'Bilangan Tenaga Buruh Bekerja')
```

## Decomposing employed labor force

```{r}
plot(decompose(lf_emp))
```

```{r}
plot(decompose(lf_emp)$trend, main = 'Trend Tenaga Buruh Bekerja di Malaysia',
     xlab = 'Tahun', ylab = 'Tenaga Buruh Bekerja')
```

```{r}
plot(decompose(lf_emp)$seasonal, 
     main = 'Corak Bermusim Tenaga Buruh Bekerja di Malaysia',
     xlab = 'Tahun', ylab = 'Tenaga Buruh Bekerja', xlim = c(2010, 2011))
```

```{r}
plot(decompose(lf_emp)$random, 
     main = 'Corak Rawak Tenaga Buruh Bekerja di Malaysia',
     xlab = 'Tahun', ylab = 'Tenaga Buruh Bekerja')
```

## Meramal Tenaga Buruh yang bekerja di Malaysia

```{r}
plot(forecast(auto.arima(lf_emp)), 
     main = 'Tenaga Buruh Bekerja Malaysia: Ramalan 2025-2026', 
     xlab = 'Tahun', ylab = 'Bilangan Tenaga Buruh Bekerja')
```

# Unemployed labor force

```{r}
lf_unemp = ts(lf[,4], start = c(2010,1), frequency = 4)

plot.ts(lf_unemp, main = 'Tenaga Buruh yang Tidak Bekerja di Malaysia', 
        xlab = 'Tahun', ylab = 'Bilangan Tenaga Buruh yang Tidak Bekerja')
```

```{r}
plot(decompose(lf_unemp)$trend, 
     main = 'Trend Tenaga Buruh Menganggur di Malaysia',
     xlab = 'Tahun', ylab = 'Tenaga Buruh Menganggur')
```

```{r}
plot(decompose(lf_unemp)$seasonal, 
     main = 'Corak Bermusim Buruh Menganggur di Malaysia',
     xlab = 'Tahun', ylab = 'Tenaga Buruh Menganggur', xlim = c(2010,2011))
```

## Decomposing unemployed labor force

```{r}
plot(decompose(lf_unemp))
```

## Meramal Tenaga Buruh yang tidak bekerja di Malaysia

```{r}
plot(forecast(auto.arima(lf_unemp)), 
     main = 'Tenaga Buruh Menganggur di Malaysia: Ramalan 2025-2026', 
     xlab = 'Tahun', ylab = 'Bilangan Tenaga Buruh Menganggur')
```

# Unemployment rate

```{r}
u_rate = ts(lf[,8], start = c(2010,1), frequency = 4)

plot.ts(u_rate, main = 'Kadar Pengangguran di Malaysia', 
        xlab = 'Tahun', ylab = 'Kadar Pengagguran')
```

## Decomposing unemployment rate

```{r}
plot(decompose(u_rate))
```

```{r}
plot(decompose(u_rate)$trend, main = 'Trend Kadar Pengangguran di Malaysia',
     xlab = 'Tahun', ylab = 'Kadar Pengangguran')
```

```{r}
plot(decompose(u_rate)$seasonal, main = 'Corak Bermusim Kadar Pengangguran di Malaysia',
     xlab = 'Tahun', ylab = 'Kadar Pengangguran', xlim = c(2010,2011))
```

## Meramal kadar pengagguran di Malaysia

```{r}
plot(forecast(auto.arima(u_rate)), 
     main = 'Kadar Pengagguran Malaysia: Ramalan 2025-2026', 
     xlab = 'Tahun', ylab = 'Kadar Pengangguran')
```

# Outside labor force

```{r}
outside_lf = ts(lf[,5], start = c(2010,1), frequency = 4)

plot.ts(outside_lf, main = 'Tenaga Buruh Luar di Malaysia', 
        xlab = 'Tahun', ylab = 'Bilangan Tenaga Buruh Luar')
```

## Decomposing outside labor force

```{r}
plot(decompose(outside_lf))
```

```{r}
plot(decompose(outside_lf)$trend, 
     main = 'Trend Luar Tenaga Buruh di Malaysia',
     xlab = 'Tahun', ylab = 'Luar Tenaga Buruh')
```

## Meramal tenaga buruh luar di Malaysia

```{r}
plot(forecast(auto.arima(outside_lf)), 
     main = 'Tenaga Buruh Luar di Malaysia: Ramalan 2025-2026', 
     xlab = 'Tahun', ylab = 'Bilangan Tenaga Buruh Luar')
```

# Significant Findings:

-   Structural break in 2020: Unemployment rate jumped from 3.5% to 5.1% due to COVID-19

<!-- -->

-   Strong recovery trend: Post-COVID recovery bringing unemployment rates back to pre-pandemic levels

<!-- -->

-   Seasonal patterns: Slight seasonal variations in employment, typically higher in Q3 and Q4

<!-- -->

-   Long-term trend: Overall increasing labor force participation rate from 63.5% (2010) to 70.5% (2024)

<!-- -->

-   Recent stability: Unemployment rate has stabilized around 3.3% in recent quarters

The analysis shows that Malaysia's labor market has been resilient, showing strong recovery post-COVID and maintaining stable unemployment rates while increasing labor force participation.
