---
title: "Tenaga Buruh"
output: 
  pdf_document:
    toc: true
    toc_depth: 6
geometry: margin=0.5cm
---

# Load library

```{r}
library(arrow)
library(forecast)
lf = read_parquet("https://storage.dosm.gov.my/labour/lfs_qtr.parquet")
```

# Labor force

## Decomposing labor force

```{r}
lf_ts = ts(lf[,2], start = c(2010,1), frequency = 4)

plot(decompose(lf_ts))
```

## Forecasting labor force

```{r}
par(mfrow=c(1,2))
plot(forecast(auto.arima(lf_ts)))
plot(forecast(tbats(lf_ts)))
```

# Employed labor force

## Decomposing employed labor force

```{r}
lf_emp = ts(lf[,3], start = c(2010,1), frequency = 4)

plot(decompose(lf_emp))
```

## Forecasting employed labor force

```{r}
par(mfrow=c(1,2))
plot(forecast(auto.arima(lf_emp)))
plot(forecast(tbats(lf_emp)))
```

# Unemployed labor force

## Decomposing unemployed labor force

```{r}
lf_unemp = ts(lf[,4], start = c(2010,1), frequency = 4)

plot(decompose(lf_unemp))
```

## Forecasting unemployed labor force

```{r}
par(mfrow=c(1,2))
plot(forecast(auto.arima(lf_unemp)))
plot(forecast(tbats(lf_unemp)))
```

# Unemployment rate

## Decomposing unemployment rate

```{r}
u_rate = ts(lf[,8], start = c(2010,1), frequency = 4)

plot(decompose(u_rate))
```

## Forecasting unemployment rate

```{r}
par(mfrow=c(1,2))
plot(forecast(auto.arima(u_rate)))
plot(forecast(tbats(u_rate)))
```

# Outside labor force

## Decomposing outside labor force

```{r}
outside = ts(lf[,5], start = c(2010,1), frequency = 4)

plot(decompose(outside))
```

## Forecasting outside labor force

```{r}
par(mfrow=c(1,2))
plot(forecast(auto.arima(outside)))
plot(forecast(tbats(outside)))
```
